---
title: "R Notebook"
output: html_notebook
---

#Installing and loading the nescessary packages
```{r}
library(XML)
library(methods)
library(sqldf)
library(data.table)
```

#Reading the XML file into R
```{r}
ReadXML <- xmlParse("/Users/mahvashmaghrabi/Downloads/Books-v4.xml")
print(ReadXML)
```

#Q1) What are the titles of all books selling for more than $50? List the titles.

```{r}
titles <- xpathSApply(ReadXML,'//book[price<50.00]/title',xmlValue)
print(titles)
```

#Q2) Who is the author of the book with the title "Oberon's Revenge"?

```{r}
author <- xpathSApply(ReadXML,'/catalog/book[ title = "Oberon\'s Revenge"]/author/text()',xmlValue)
print(author)
```

#Q3) What are the titles of the books that sell for the lowest price?

```{r}
lowCost <- xpathSApply(ReadXML,'//book[price=min(//book/price)]/title',xmlValue)
print(lowCost)
```

#Q4)List the title and prices of all books in the "Computer" genre.

```{r}
compGenre <- xpathSApply(ReadXML,'//book[genre = "Computer"]/title/text() | //book[genre = "Computer"]/price/text()',xmlValue)
print(compGenre)
```

#Q5)How many books cost more than $50?

```{r}
cost <- xpathSApply(ReadXML,'count(//book[price > 50])',xmlValue)
print(cost)
```

# Converting the XML file to a DataFrame
```{r}
xmldata <- xmlToDataFrame("/Users/mahvashmaghrabi/Downloads/Books-v4.xml")
print(xmldata)
```
# Converting the dataframe into a data table
```{r}
catalogTable <- setDT(xmldata)
```

# Creating and connecting to the new database
```{r}
mydb <- dbConnect(drv = RSQLite::SQLite(), dbname = "booksdb")
```

# Writing the data to the database table created
```{r}
dbWriteTable(conn = mydb, name="catalogTable", value = catalogTable, overwrite = T)
```

#Select query from the new database created
```{sql connection = mydb}
SELECT * FROM catalogTable

```
#Part2
#How many books are in each genre and what is the average price for the books in each genre? List the genre, number of books, and average price.

```{sql connection = mydb}
SELECT genre AS GENRE,count(*) AS "NUMBER OF BOOKS",avg(price) AS 
"AVERAGE PRICE" FROM catalogTable GROUP BY genre;
```

