---
title: "R Notebook"
output: html_notebook
---

# Installing and loading the necessary packages
```{r}
library(XML)
library(methods)
```

# Reading the XML file into R
```{r}
ReadXML <- xmlParse(file = "/Users/mahvashmaghrabi/Desktop/XMLFile.xml")
```

# Printing the XML file
```{r}
print(ReadXML)
```

# Loading the necessary package
```{r}
library(xml2)
```

# Choosing the XML file that needs to be validated
```{r}
XMLfile <- xml2::read_xml(file.choose(""))

```

# Choosing the XML schema against which the XML file will be validated
```{r}
XMLschema <- xml2::read_xml(file.choose())
```

# Validating the XML file and checking for errors
```{r}
xml2::xml_validate(XMLfile,XMLschema)
validation <- xml2::xml_validate(XMLfile,XMLschema)
```

# Printing the errors if any
```{r}
print(attr(validation,"errors"))
```

# XPath expression to find the number of visits to the page 120
```{r}
Xpath <- "count(//log//visit[@Id=20])"
Visits <- xpathApply(ReadXML, Xpath, xmlValue)
print(paste("Number of visits to the page 120 :", Visits))
```

